---
title: "History of new commanders"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{history-of-cmders}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
library(cowplot)
library(gganimate)
# devtools::install_github("hadley/emo")
library(emo)
library(scryr)
library(mtggplot)
```

How has the number of commanders available to each color grown over time? 
Technically, we could just use `scry_cards("is:commander")`, but we using a more
complicated search because we want to include at least one card from each set 
that added cards to the format. That will make it easier to count new commanders
by set, even when that number is zero. We're also exluding a few promotional 
sets (book promos, 
[Nalathni Dragon](https://scryfall.com/card/pdrc/1/nalathni-dragon), and Game 
Night).
```{r}
legal_cmder_search <- scry_cards("(t:creature or o:'can be your commander') not:reprint f:commander -set:pred -set:gnt -set:phpr")

# writing to local file to avoid having to re-run query
# write_rds(legal_cmder_search, here::here("vignettes", "legal_cmder_search.rds"))
```

